### 2024-05-20 manuscript.Rmd
update manuscript and prepare for submission

---
title             : "Digital video platforms and phylogenetic modelling reveal parrot tool use is not rare"
shorttitle        : "Parrot tool use"

author: 
  - name          : "Amalia P. M. Bastos"
    affiliation   : "†,1,2"
  - name          : "Scott Claessens"
    affiliation   : "*†,2"
  - name          : "Ximena J. Nelson"
    affiliation   : "3"
  - name          : "David Welch"
    affiliation   : "4"
  - name          : "Quentin D. Atkinson"
    affiliation   : "‡,2"
  - name          : "Alex H. Taylor"
    affiliation   : "‡,2,3,5,6"

affiliation:
  - id            : "1"
    institution   : "\\footnotesize Department of Psychological & Brain Sciences, Johns Hopkins University, Baltimore, MD, United States"
  - id            : "2"
    institution   : "\\footnotesize School of Psychology, University of Auckland, Auckland, New Zealand"
  - id            : "3"
    institution   : "\\footnotesize School of Biological Sciences, University of Canterbury, Christchurch, New Zealand"
  - id            : "4"
    institution   : "\\footnotesize School of Computer Science, University of Auckland, Auckland, New Zealand"
  - id            : "5"
    institution   : "\\footnotesize ICREA, Pg. Lluís Companys 23, Barcelona, Spain"
  - id            : "6"
    institution   : "\\footnotesize Institute of Neuroscience, Universitat Autònoma de Barcelona, Barcelona, Spain"

abstract: |
 Putatively rare behaviours, such as tool use, are challenging to study because absence of evidence can arise either from a species’ inability to produce the behaviour or from insufficient research effort. Here, we tackle this challenge by combining data from digital platforms and phylogenetic modelling to estimate actual rates of tool use in parrots. Videos on a social media platform revealed novel instances of tool use in 17 parrot species, more than doubling the confirmed number of tool-using parrot species from 11 (3%) to 28 (7%). Phylogenetic modelling ranked additional species that are most likely to be unobserved tool users, suggesting that between 11% and 17% of extant parrot species may be tool users. These discoveries have implications for inferences about the evolutionary drivers and origins of tool use in parrots, revealing associations with relative brain size and feeding generalism and indicating several genera where tool use was likely an ancestral trait. Overall, our findings challenge the assumption that current sampling effort captures the full distribution of putatively rare animal behaviours. Combining our sampling and analysis methods offers a fruitful approach for investigating the distribution, drivers, and origins of other rare behaviours.

keywords          : "tool use; technical intelligence; parrots; digital video platforms; phylogenetic modelling"
bibliography      : ["r-references.bib"]
floatsintext      : yes
figurelist        : no
figsintext        : yes
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
numbersections    : false
documentclass     : "apa6"
classoption       : "man, donotrepeattitle"
output            : papaja::apa6_pdf
header-includes: 
  - \usepackage{array}
  - \usepackage{caption}
  - \usepackage{graphicx}
  - \usepackage{siunitx}
  - \usepackage[normalem]{ulem}
  - \usepackage{colortbl}
  - \usepackage{multirow}
  - \usepackage{hhline}
  - \usepackage{calc}
  - \usepackage{tabularx}
  - \usepackage{threeparttable}
  - \usepackage{wrapfig}
  - \usepackage{adjustbox}
  - \usepackage{hyperref}
  - \usepackage{setspace}
  - \raggedbottom
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
  - \AtBeginEnvironment{tablenotes}{\doublespacing}
  - \captionsetup[table]{font={stretch=1,small}}
  - \captionsetup[figure]{font={stretch=1,small}}
  - \nolinenumbers
  - \note{\raggedright † These authors contributed equally to this work \par ‡ These authors contributed equally to this work \par * Correspondence concerning this article should be addressed to Scott Claessens, Level 2, Building 302, 23 Symonds Street, Auckland, New Zealand. \text{E-mail:} \href{scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}} \par This working paper has not yet been peer-reviewed.}
appendix          : "appendix.Rmd"
csl               : current-biology.csl
---

\newpage
\nolinenumbers

# Introduction

Our understanding of the evolution of animal behaviour is built on the 
assumption that we have access to sufficient data [@MacLean2012; @Lefebvre2011; 
@Logan2018]. However, this is not always the case. Data on behaviours that are 
rare, fleeting, or otherwise difficult to observe are likely to be patchy and 
incomplete [@Healy2007; @Powell2017]. Among species for which such behaviours 
have not been observed, it can be difficult to differentiate between cases in 
which the species is truly incapable of producing the behaviour and cases in 
which the species is capable of producing the behaviour but the behaviour has 
not yet been observed. Such a distinction can be critical for drawing 
conclusions about the rarity and evolution of the behaviour in question.

Comparative work on the evolution of tool use is a paradigmatic example of this 
issue. The initial discoveries of tool use in chimpanzees [@Goodall1964], 
birds [@Hunt1996], dolphins [@Smolker1997], and octopuses [@Finn2009] occurred 
decades after significant advances on other more easily measurable aspects of 
their biology. Since then, scholars have proposed a clear operational 
definition of tool use applicable to all species --- the manipulation of an 
unattached object as an extension of the animal's body to achieve a 
goal [@Shumaker2011] --- and have used the distribution of species meeting this 
definition to make various claims about the evolutionary drivers of tool use 
behaviours. For example, based on the observation that bird species with 
reported tool use tended to have larger brains, researchers have identified 
higher relative brain size as a likely precondition for tool using 
capabilities [@Lefebvre2002;@Lefebvre1997;@Lefebvre2004]. These researchers 
argue that larger brains are better able to integrate visual and somatosensory 
information when innovating novel behaviours, such as tool use, in changing 
environments [@Sol2005;@Sol2009]. Similarly, researchers have used existing 
reports of tool use in birds to debate the roles of generalist versus 
specialist feeding strategies in driving the evolution of tool use, with some 
arguing that feeding generalists require technical innovations to expand their 
dietary niche [@Ducatez2015;@Overington2011;@Sol2005] and others arguing that 
feeding specialists require technical innovations for extractive foraging of 
specific foods [@HenkeVonDerMalsburg2020;@MettkeHofmann2014].

However, before we can make claims of this kind, we need to know whether current 
research effort in the literature is sufficient for robust conclusions to be 
drawn about the evolution of tool use. In fact, evidence suggests that research 
effort is often systematically biased towards particular taxonomic groups, parts 
of the world that are easy to access, and species with life history traits that 
make them easier to study, such as larger distribution ranges and population 
sizes [@Ducatez2014]. This is a crucial limitation because insufficient 
observation may lead researchers to miss true instances of tool behaviours and 
thus draw premature conclusions about the evolutionary drivers and origins of 
tool use. While researchers have attempted to deal with this problem by 
controlling for the number of scientific papers published on different species, 
previous work has not yet attempted to quantify and explicitly model the 
relationship between actual tool-using behaviour and what is reported in the 
scientific literature. If more tool-using species exist than previously thought, 
this could have important implications for theories of the evolutionary drivers 
and origins of tool use and for our understanding of how rare this behaviour 
actually is.

One potentially powerful method for quantifying actual rates of rare animal 
behaviours is by collating evidence of the behaviours from digital video 
platforms [@Nelson2013]. With millions of videos posted every day, digital 
platforms like YouTube offer a valuable source of data on animal behaviour. 
Digital video platforms have already been successfully used to uncover a variety
of rare animal behaviours, including interspecies play in dogs [@Nelson2013], 
novel problem-solving behaviours in horses [@Krueger2019], and death-related 
behaviours in Asian elephants [@Pokharel2022]. By casting the net wider than the 
scientific literature, digital video platforms can provide an indication of the 
tool-using species that the literature might be missing.

Even after collating evidence from digital video platforms, some tool-users 
could _still_ remain unobserved. One principled framework for identifying these
unobserved species is to specify a causal model of the process that generates
the observed data. We propose one such causal model in Figure 
\@ref(fig:plotDAG). In this model, we assume that the presence of tool use in 
the scientific literature (or on digital video platforms) is caused by both 
unobserved tool use capabilities and the number of published studies (or the 
number of videos) for any given species. Tool users are more likely to be 
observed if they are well studied, but understudied tool users may go 
undetected. Furthermore, based on existing theories of the evolution of tool use
[@Ducatez2015;@HenkeVonDerMalsburg2020;@Lefebvre1997;@Lefebvre2002;
@Lefebvre2004;@MettkeHofmann2014;@Overington2011;@Sol2005;@Sol2009], we propose
that the unobserved tool use capabilities are additionally caused by relative 
brain size, feeding strategy, and shared phylogenetic ancestry. Expressing this
causal model as a statistical model can suggest further species which are likely
to be unobserved tool-users and, simultaneously, test existing theories of the 
evolutionary drivers of tool use without incorrectly assuming that absence of 
evidence is evidence of absence.

(ref:plotDAGCaption) _Causal model of observed tool use._ Directed acyclic graph 
of the causal relationships between observed tool use and other variables. 
Available scientific data on tool use is caused both by unobserved tool use 
presence and scientific research effort (i.e., number of publications). 
Available video data on tool use is caused both by unobserved tool use 
presence and video research effort (i.e., number of videos). According to 
theory, unobserved tool use presence should be caused by relative brain size 
(encephalisation quotient) and feeding strategy (generalist vs. specialist). 
These variables all share unobserved confounds generated by shared phylogenetic
history. Grey circles indicate unobserved variables.

```{r plotDAG, echo=F, fig.width=6, fig.height=3.6, fig.cap="(ref:plotDAGCaption)"}
tar_read(plotDAG)
```

Here, we apply these approaches to tool use in the parrot order. We focus on 
tool use in parrots for a number of reasons. First, the scientific literature 
suggests that only a small proportion of extant parrot species (11 out of 398;
3%) use tools [@Auersperg2009;@Auersperg2010;@Auersperg2011a;@Auersperg2011b;
@Auersperg2012;@Bastos2021;@BentleyCondit2010;@Borsari2005;@Goodman2018;
@Heinsohn2017;@Janzen1976;@Lambert2015;@Lefebvre2002;@OHara2021]. Parrot tool 
use thus provides an ideal test case for examining how robust sampling is in the 
scientific literature. Second, parrots are highly popular as pets. Over 70% of 
all extant parrot species are bred in the aviculture industry and kept as pets 
worldwide [@Anderson2003;@Carrete2008;@Drews2001;@Kelly2014;@Li2014;@Su2015;
@Young2012], enabling us to leverage the power of digital video platforms to 
search for evidence of tool use [@Nelson2013]. Third, detailed data on relative
brain sizes [@Flammer2001;@Iwaniuk2005;@Mazengenya2018;@Olkowicz2016;
@Schuck2008;@Silva2017], feeding strategies [@Wilman2016], and shared ancestry
[@Jetz2012] exist for parrots, allowing us to fit the statistical model implied
by Figure \@ref(fig:plotDAG) to the entire parrot order.

We first present the results from our video survey, in which we collated videos
of tool use in parrots from a digital video platform. This survey reveals a 
number of previously unidentified tool-using parrot species, which we map onto 
the phylogeny of the parrot order. We then describe our statistical model in 
more detail, and use it to (_i_) rank further parrot species that are likely 
unobserved tool users and (_ii_) re-examine key hypotheses regarding the 
evolutionary drivers and origins of tool use in parrots.

# Results

## Digital video platform reveals tool use in additional parrot species

We surveyed the digital media platform YouTube for video evidence of tool use in 
parrots (see Methods for detailed search criteria). In our search, we used the 
standard criteria for identifying tool use in the literature, defining "true" 
tool use behaviour as the manipulation of an unattached object as an extension 
of the animal's body to achieve a goal [@Shumaker2011], while "borderline" tool 
use involved the use of an object that was still attached to a substrate
[@Seed2010].

```{r echo=F}
tar_load(crowdData)
trueCrowdData <- filter(crowdData, ToolUse == "Y")
bordCrowdData <- filter(crowdData, ToolUse == "B")
```

In total, we found `r nrow(crowdData)` videos of 
`r crowdData %>% pull(IndividualID) %>% unique() %>% length()` individuals from 
`r crowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
parrot species performing behaviours that met the definition of either true tool 
use (`r nrow(trueCrowdData)` videos of 
`r trueCrowdData %>% pull(IndividualID) %>% unique() %>% length()` 
individuals from 
`r trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
species) or borderline tool use (`r nrow(bordCrowdData)` videos of 
`r bordCrowdData %>% pull(IndividualID) %>% unique() %>% length()` 
individuals from 
`r bordCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
species). All videos featured pet parrots in captive settings. In 
`r crowdData %>% pull(HumanInfluence) %>% str_starts("No") %>% sum()` of these 
videos, owners did not appear to interact with the subjects. In 
`r crowdData %>% pull(HumanInfluence) %>% str_starts("Yes") %>% sum()` videos, 
there was potential human interaction, either from the owners being in close physical 
contact with the bird (e.g., bird perching on hand), talking to the bird, or 
handing it the tool (which occurred in only two videos). We could not establish 
the degree of human interaction in the remaining 
`r crowdData %>% pull(HumanInfluence) %>% str_starts("Unknown") %>% sum()` 
videos, as sound had been removed or was substituted by music. None of the 
videos featured owners directly rewarding tool use behaviours with food. All 
borderline tool use cases were excluded from further analyses.

Of the `r trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
parrot species performing true tool use, 
`r trueCrowdData %>% filter(Species != "Unknown") %>% group_by(Species, IndividualID) %>% summarise(n = n(), .groups = "drop") %>% group_by(Species) %>% summarise(check = n() > 1, .groups = "drop") %>% pull(check) %>% sum()` 
were represented in our video survey by two or more individuals over multiple 
independent observations. True tool use always involved the subject using an 
object for self-scratching (`r trueCrowdData %>% mutate(BodyPartScratched = str_detect(BodyPartScratched, "Head/Neck")) %>% pull(BodyPartScratched) %>% sum()` 
videos involved scratching the head and/or neck). The most common tool (`r sum(pull(trueCrowdData, EvaluatedTool) == "Feather")`
videos) was a moulted feather. Human-made objects (e.g., pens, spoons, pieces of 
wood, cardboard) were also common.

According to YouTube video descriptions and owner comments, 
`r trueCrowdData %>% group_by(IndividualID, Sex) %>% summarise(.groups = "drop") %>% pull(Sex) %>% (function(x) sum(x == "Male"))` 
of the individuals performing true tool use were males and 
`r trueCrowdData %>% group_by(IndividualID, Sex) %>% summarise(.groups = "drop") %>% pull(Sex) %>% (function(x) sum(x == "Female"))` 
were females. No sex information was provided for the remaining 
`r trueCrowdData %>% group_by(IndividualID, Sex) %>% summarise(.groups = "drop") %>% pull(Sex) %>% (function(x) sum(x == "Unknown"))` 
individuals. As owners provided no information on whether sex had been 
established through genetic testing, and sexual dimorphism in parrots is 
rare [@Bercovitz1987;@Hoyo2011], we could not typically ascertain if descriptions 
were accurate. It is unclear if the disproportionately large number of males in 
the sample is a consequence of owners more readily assuming their parrots are 
male when they have not been genetically tested, owners being more likely to own 
or film male parrots, or male parrots exhibiting more true tool use behaviours 
than female parrots.

Figure \@ref(fig:plotPhylo1) maps the findings from the video survey onto a 
maximum clade credibility phylogeny for the parrot order, plotted alongside 
species previously identified in the scientific literature. Before the video 
survey, 11 parrot species (3%) had been identified as tool users in the 
scientific literature. Across our video survey, we observed true tool use in 
`r trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
species, 5 of which overlapped with the scientific literature and 
`r (trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()) - 5` 
of which were novel species. All of the species identified in the video survey 
were cockatoos (_Cacatuidae_), Old World parrots (_Psittacinae_), or neotropical 
parrots (_Arinae_). The most common species in our survey, accounting for 
`r crowdData %>% filter(Species == "Pyrrhura molinae") %>% nrow()` videos from 
`r crowdData %>% filter(Species == "Pyrrhura molinae") %>% pull(IndividualID) %>% unique() %>% length()` 
individuals, was the green-cheeked conure (_Pyrrhura molinae_). In accordance 
with the scientific literature, the video survey found no evidence of tool use 
in any species of Psittaculidae, despite this family containing some of the most 
commonly kept pet species, including lovebirds, lorikeets, and Asian parakeet 
species. Combining both the video survey and the scientific literature, we can 
thus identify 28 tool-using parrot species overall (7%), compared to the 11 
previously reported.

(ref:plotPhylo1Caption) _Results of video survey and phylogenetic survival cure
modelling mapped onto a maximum clade credibility phylogeny of the parrot order._
Orange points in the inner ring indicate species observed in the video survey,
with point size scaled by the number of videos for each species (note that three
species observed only in the video survey are not present in the phylogeny due
to a lack of genomic data: _Psittacara erythrogenys_, _Psittacus timneh_, and 
_Aratinga nenday_). Blue points in the outer ring indicate species observed in
the scientific literature, with point size scaled by the number of papers 
published on each species. Yellow species names indicate the top ten most likely
tool-using species from our phylogenetic survival cure model which were not 
observed in the scientific literature or the video survey. Total _n_ = 
`r nrow(tar_read(d))` species.

```{r plotPhylo1, echo=F, fig.width=17, fig.height=15, fig.cap="(ref:plotPhylo1Caption)"}
tar_read(plotPhylo1)
```

The identification of new tool-using species in our video survey increases the 
extent to which phylogeny can explain the distribution of tool use in the parrot 
order. We estimated phylogenetic signal (Pagel's $\lambda$) of tool use using 
both the pre-video-survey and post-video-survey data. Pagel's $\lambda$ varies 
between 0 and 1, where 0 implies that the distribution of a trait across species 
is unexplained by phylogenetic relatedness and 1 implies that the distribution 
of a trait across species is fully explained by phylogeny. Using the evidence of 
tool use from the scientific literature alone (pre-video-survey data; 11 
tool-using species), we estimated an average posterior Pagel's $\lambda$ of 
`r printnum(mean(tar_read(phySignal1)))` (95% credible interval [`r printnum(quantile(tar_read(phySignal1), 0.025))` 
`r printnum(quantile(tar_read(phySignal1), 0.975))`]; total _n_ = 
`r nrow(tar_read(d))` species). This estimate was moderate-to-strong, but highly 
uncertain. In comparison, combining the evidence from both the literature and 
the video survey (post-video-survey data; 28 tool-using species) resulted in a 
stronger and more certain estimate of phylogenetic signal. With these data, we 
estimated Pagel's $\lambda$ = `r printnum(mean(tar_read(phySignal2)))` (95% CI 
[`r printnum(quantile(tar_read(phySignal2), 0.025))` 
`r printnum(quantile(tar_read(phySignal2), 0.975))`]; total _n_ = 
`r nrow(tar_read(d))` species). Thus, the results of our video survey increase 
the extent to which the distribution of tool use across parrot species can be 
explained by shared phylogenetic ancestry. This suggests that we can potentially 
use phylogenetic information, along with other variables, to identify further 
tool-using parrot species that may remain undetected.

## Phylogenetic survival cure modelling identifies further candidate tool users

In addition to the 28 tool-using species identified in the literature and our 
video survey, we fitted a Bayesian phylogenetic survival cure model to rank 
further species that are likely to be undetected tool-users (i.e., tool-using 
species with no tool use reported in the literature or in videos).

Survival cure models [@Amico2018], also known as split population 
models [@Schmidt1989], are used to analyse the time to some event of interest 
with the added assumption that a certain proportion of the population will never 
experience the event, no matter how long they are measured for. These models 
have been used to analyse a variety of right-censored outcomes, from cancer 
relapse [@Amico2018] to criminal recidivism [@Schmidt1989]. The data are 
right-censored because some individuals will have experienced the event when 
they are measured (e.g., disease onset, return to prison) while others will have 
not experienced the event. For those who have not experienced the event, this 
may be because (_i_) the event has not happened to them yet or (_ii_) the event 
will _never_ happen to them. Survival cure models treat these two processes 
separately.

Our tool use problem has the same features. We are modelling a time-to-event; 
specifically, the amount of "time" (i.e., observation opportunities measured as 
the number of published papers or videos) until tool use is identified. This is
right-censored data because many species will not have had tool use identified
when we measure them. Moreover, we can assume that a certain proportion of the
population will never experience the event -- that is, they are a non-tool-user
and so we will never identify tool use no matter how long we measure them for.

In our model, we infer the tool-using status of each species by allowing each 
species to have their own probability of being a non-tool-user. Following our 
causal model (Figure \@ref(fig:plotDAG)), we predict these probabilities based 
on feeding strategy, encephalisation quotient, and phylogenetic history (see 
Methods for full model). The model additionally takes research effort into 
account by allowing that, among species for which tool use is unobserved, all 
else being equal those with fewer published papers and fewer video search hits 
have a higher probability of being undetected tool users (Supplementary Figure 
\@ref(fig:plotSurvCure3)).

```{r echo=FALSE}
tar_load(plotSurvCure1)
```

We found that this phylogenetic survival cure model was able to adequately 
distinguish between species with and without evidence for tool use, with an 
area-under-the-curve classification statistic of 
`r printnum(tar_read(auc), digits = 2)` (Supplementary Figure 
\@ref(fig:plotSurvCure9)). To further estimate the accuracy of the model's 
predictions, we also used a leave-one-species-out approach with known tool 
users. For each of the 25 tool-using species that were represented on the 
phylogeny and for which we had brain size and genomic data (we lacked data for 
three tool-using species), we fitted the model to a modified dataset which set 
tool use to be absent for the target species in both the scientific literature 
and the video survey. Across `r length(tar_read(cv))` cross-validation models, 
`r sum(tar_read(cv))` models (`r printnum(mean(tar_read(cv))*100, digits = 0)`%) 
continued to predict the target species as having a median posterior probability 
of tool use that was within the range of all other tool users. This 
classification rate was greater than the baseline classification rate of 
`r printnum(mean(plotSurvCure1$data %>% filter(ToolUse == "Absent") %>% pull(med) > 0.16478274) * 100, digits = 0)`% 
for species without evidence of tool use in the full model
(`r sum(plotSurvCure1$data %>% filter(ToolUse == "Absent") %>% pull(med) > 0.16478274)` 
of `r nrow(plotSurvCure1$data %>% filter(ToolUse == "Absent"))` species without 
evidence of tool use had a median posterior probability of tool use that was 
within the range of the tool-using species). Together, the area-under-the-curve 
statistic and the leave-one-species-out approach suggest that the model is able 
to adequately classify known tool users, with some error.

Figure \@ref(fig:plotSurvCure1) visualises the ranked posterior probabilities of 
tool use from the phylogenetic survival cure model for all parrot species. As 
expected, the known tool users are ranked towards the top of this list. However, 
several "tool use absent" species also rank highly on the list, despite not 
being identified as tool users in the scientific literature or in our video 
survey. In fact, according to the model, the most likely tool user is a species 
for which tool use is unobserved in our data: the blue-eyed cockatoo (_Cacatua 
opthalmica_). This species is endemic to Papua New Guinea and is relatively 
understudied, with only `r tar_read(d)$EX.Pre.Tool.Papers[165]` published papers 
and `r tar_read(d)$SearchHits[165]` video search hits, which is fewer than the 
model expects are necessary to discover tool use when it is present (Figure 
\@ref(fig:plotSurvCure2)). This species is also found in the _Cacatua_ genus, a 
clade containing several known tool users. This prediction makes sense given the 
high phylogenetic signal for tool use reported by the model (Supplementary 
Figures \@ref(fig:plotSurvCure5) and \@ref(fig:plotSurvCure6)). Beyond the 
blue-eyed cockatoo, other highly ranked species without observed evidence of 
tool use are the Meyer's parrot (_Poicephalus meyeri_), the golden parakeet 
(_Guaruba guarouba_), the long-billed corella (_Cacatua tenuirostris_), the 
Solomons cockatoo (_Cacatua ducorpsii_), the red-fronted parrot (_Poicephalus 
gulielmi_), the Cape parrot (_Poicephalus robustus_), the yellow-eared parrot 
(_Ognorhynchus icterotis_), the red-vented cockatoo (_Cacatua haematuropygia_), 
and the gang-gang cockatoo (_Callocephalon fimbriatum_). Figure 
\@ref(fig:plotPhylo1) plots these species on the parrot phylogeny, using the top 
ten highest ranked species without observed evidence of tool use as an arbitrary 
cutoff for visualisation purposes.

(ref:plotSurvCure1Caption) _Posterior predicted probabilities of tool use for 
each species from our phylogenetic survival cure model._ Points are posterior 
medians and lines are 50% and 95% credible intervals. Total _n_ = 
`r nrow(tar_read(d))` species.

```{r plotSurvCure1, echo=F, fig.width=7, fig.height=10, fig.cap="(ref:plotSurvCure1Caption)"}
tar_read(plotSurvCure1)
```

(ref:plotSurvCure2Caption) _Expected number of published papers and videos until 
tool use discovery, according to the survival component of the phylogenetic 
survival cure model._ Densities are full posterior distributions, points are 
posterior medians, and lines are 50% and 95% credible intervals.

```{r plotSurvCure2, echo=F, fig.width=7, fig.height=3, fig.cap="(ref:plotSurvCure2Caption)"}
tar_read(plotSurvCure2)
```

```{r echo=F}
tar_load(sumProbs)
```

The posterior probabilities shown in Figure \@ref(fig:plotSurvCure1) are 
estimated with uncertainty, so it is difficult to "identify" any particular 
species as an undetected tool user. Nevertheless, taking the sum of all the 
posterior probabilities for the 
`r nrow(plotSurvCure1$data %>% filter(ToolUse == "Absent"))` species without 
observed evidence of tool use, we can estimate that around 
`r printnum(median(sumProbs), digits = 0)` of those species are likely to be 
undetected tool users (median sum of probabilities = 
`r printnum(median(sumProbs), digits = 2)`, 95% CI
[`r printnum(quantile(sumProbs, 0.025), digits = 2)` 
`r printnum(quantile(sumProbs, 0.975), digits = 2)`]). When combined with the 
species known to use tools, this implies that between 11% and 17% of extant 
parrot species may be tool users.

## Implications for the evolutionary drivers and origins of tool use

```{r echo=FALSE}
tar_load(postFull)
diff <- postFull$aP[,2] - postFull$aP[,1]
```

The predicted probabilities from our phylogenetic survival cure model have 
implications for inferences about the evolutionary drivers and origins of tool 
use in the parrot order. Regarding the drivers of tool use hypothesised in 
Figure \@ref(fig:plotDAG), the phylogenetic survival cure model revealed that 
encephalisation quotient strongly positively predicted the probability of tool 
use (median posterior log odds slope = `r printnum(median(-postFull$bEQ))`, 95% 
CI [`r printnum(quantile(-postFull$bEQ, 0.025))` 
`r printnum(quantile(-postFull$bEQ, 0.975))`]; total _n_ = `r nrow(tar_read(d))` 
species; Figure \@ref(fig:plotSurvCure4)). This helps explain the ranking in 
Figure \@ref(fig:plotSurvCure1): the blue-eyed cockatoo has the largest relative 
brain size in the dataset. We also found that feeding generalist species were 
slightly more likely to be tool users, though the posterior difference between 
generalists and specialists was quite uncertain (median posterior log odds 
difference = `r printnum(median(diff))`, 95% CI [
`r printnum(quantile(diff, 0.025))` `r printnum(quantile(diff, 0.975))`]; total 
_n_ = `r nrow(tar_read(d))` species). These results from the survival cure model 
differed from the results of models fitted to pre-video-survey and 
post-video-survey data without the survival cure component, which found no 
effect of relative brain size and no difference between feeding strategies, 
respectively (Supplementary Figure \@ref(fig:plotComparison)).

(ref:plotSurvCure4Caption) _Posterior predictions for the effects of feeding 
strategy and encephalisation quotient on the probability of tool use from the 
phylogenetic survival cure model._ In the left plot, points and lines represent 
posterior medians and 50% and 95% credible intervals, with densities 
representing full posterior distributions. In the right plot, the line and 
shaded areas represent the posterior median regression line with 50% and 95% 
credible intervals. In both plots, individual species are coloured according to 
the presence / absence of tool use in the video survey and the scientific 
literature. Total _n_ = `r nrow(tar_read(d))` species, generalist _n_ = 
`r sum(tar_read(d)$Feeding == "Generalist")` species, specialist _n_ = 
`r sum(tar_read(d)$Feeding == "Specialist")` species.

```{r plotSurvCure4, echo=F, fig.width=8, fig.height=4, fig.cap="(ref:plotSurvCure4Caption)"}
tar_read(plotSurvCure4)
```

Regarding the origins of tool use, we fitted exploratory ancestral state 
reconstruction models to the pre-video-survey data, the post-video-survey data, 
and the predicted probabilities from the phylogenetic survival cure model. The 
discoveries from our video survey and from our phylogenetic modelling increased 
the likelihood that tool use was present in the most recent common ancestor for 
several parrot genera. These include the amazon parrots native to the Americas 
(_Amazona_), the true white cockatoos and corellas found in South East Asia and 
Australasia (_Cacatua_), the kea and the kākā from New Zealand (_Nestor_), and 
the _Poicephalus_ genus native to Africa (Table \@ref(tab:tableASR); 
Supplementary Figures \@ref(fig:plotASR1) - \@ref(fig:plotASR3)). These findings 
suggest that species from each of these genera may share their tool use 
capabilities via common descent from their respective common ancestors, rather 
than via independent evolution within each genus or behavioural innovation
within species.

(ref:tableASRCaption) Estimated probabilities of tool use for most recent common 
ancestors of several parrot genera. _Probabilities estimated using exploratory 
ancestral state reconstruction models fitted to the pre-video-survey data, 
post-video-survey data, and predicted probabilities from the phylogenetic 
survival cure model. Total_ n _= `r nrow(tar_read(d))` species._

```{r tableASR, echo=F, warning=F, message=F}
apa_table(
  tar_read(tableASR),
  caption = "(ref:tableASRCaption)",
  escape = FALSE
)
```

# Discussion

Since the earliest anecdotes of parrots using tools by Wallace in the 1880s
[@Wallace1869] and more systematic anecdotal reports in the 1970s [@Goodall1971], 
only 11 parrot species (3% of all extant parrots) have been documented as tool 
users in the scientific literature. Our study collated data from a digital video
platform to identify 17 additional tool-using parrot species that are new to 
science, more than doubling the overall count to 28 species (7%). These species 
consisted of cockatoos (_Cacatuidae_), Old World parrots (_Psittacinae_), and 
neotropical parrots (_Arinae_). Beyond the video survey, the strong 
phylogenetic signal in our dataset allowed us to use phylogenetic information, 
along with other variables, to infer the unobserved probabilities of tool use 
across the parrot order. Our phylogenetic survival cure model incorporated
information on phylogenetic history, research effort, relative brain size, and 
feeding specialisation to rank parrot species that were most likely to be 
undetected tool users. The sum of probabilities from this model implied that 
between 15 and 41 of the species without observed evidence of tool use are 
likely to be undetected tool users, suggesting that the true proportion of tool
users may be as high as 17%.

These findings have a number of important implications. First, our findings show 
that current research effort in the scientific literature is insufficient to 
capture the real world occurrence of parrot tool use. If the scientific 
literature had sampled the natural world sufficiently, we would expect to see 
close correspondence between those species reported as tool users in the 
literature and those species the public have filmed performing tool use. 
Instead, we discovered a large discrepancy between these two data sources, both 
in the prevalence of tool use and the species identified. This raises the 
possibility that insufficient research effort is a general issue across the 
scientific literature, both for tool use in other groups and for other rare 
behaviours.

Second, in terms of the evolution of tool use in parrots, our study challenges a 
key assumption made in the literature to date: that only a minority of parrots 
are tool users [@Auersperg2012;@Bastos2021;@Goodman2018;@Lambert2015;@OHara2021]. 
The paucity of evidence for tool use across parrots in the literature initially 
implied that the capacity for tool use may have evolved independently in 
different parrot species. Our discovery of the widespread distribution of tool 
use across the parrot phylogeny, along with the strong phylogenetic signal in 
this expanded dataset, challenges this and suggests that, at least for some 
parrot clades, the capacity for tool use might be a homologous trait that has 
been evolutionarily conserved. Our exploratory ancestral state reconstruction 
analysis provides preliminary support for this hypothesis, revealing an 
increased probability of tool use among the most recent common ancestors for the 
_Amazona_, _Cacatua_, _Nestor_, and _Poicephalus_ genera. Even at this 
preliminary stage, our analysis therefore raises an alternative hypothesis for 
the observed tool use in _Cacatua_ [@Auersperg2012;@OHara2021] and 
_Nestor_ [@Auersperg2009;@Auersperg2010;@Auersperg2011a;@Auersperg2011b;
@Bastos2021;@Goodman2018], namely that tool behaviours have arisen due to the 
common ancestor having the capacity to use tools, rather than from independent 
evolution or behavioural innovation within species.

Third, our results support existing theories of the drivers of tool use. We 
found that encephalisation was strongly positively related to the probability of 
tool use in our phylogenetic model, supporting previous theories linking 
relative brain size to increased tool innovation in birds [@Lefebvre1997;
@Lefebvre2002;@Lefebvre2004] and primates [@Reader2002]. We acknowledge that
encephalisation quotient is not a perfect measure due to measurement error and
challenges with interpretation [@Healy2007;@Logan2018]. However, encephalisation
quotient has much greater coverage across the parrot phylogeny than more 
fine-grained measures like whole neuron count [@Sol2022] and there is no reason 
to think that measurement error would produce the consistent patterns across our
study and prior work. To understand the causal mechanisms responsible for these
relationships, we encourage further work on the specific neural correlates of 
technical intelligence in parrots, e.g., [@Alvarez2020]. In our phylogenetic 
model, we also found that tool use was somewhat more likely among feeding 
generalists compared to feeding specialists, although this difference was 
uncertain. This trend supports previous suggestions that increased cognitive 
abilities and technical innovation rates are required to expand a generalist 
species' dietary niche [@HenkeVonDerMalsburg2020;@HenkeVonDerMalsburg2021;
@Ducatez2015]. However, the trend contradicts theories linking tool use to 
dietary specialisation, whereby species eating specific foods that require 
extractive foraging have higher cognitive ability and are especially prone to 
using tools [@HenkeVonDerMalsburg2020].

All instances of tool use in our video survey met the established criteria for 
tool use in the literature: the manipulation of an unattached object as an 
extension of the animal's body to achieve a goal [@Shumaker2011]. However, a 
striking feature of the dataset is that the 17 additional species identified in
the video survey were exclusively observed using tools for self-scratching.
While parrots in the wild use tools to achieve a variety of goals, including 
extractive foraging and preening [@Goodman2018;@OHara2021], parrots in captivity
only require tools to achieve the latter goal. Despite this difference in the
type of tool use observed, self-scratching is nonetheless a major category of 
animal tool use. Self-scratching also meets the definition for a more complex 
type of embodied tool use known as 'tooling', the deliberate generation of a 
mechanical interface by using an object to manipulate another target or surface
[@Fragaszy2018]. This makes it difficult to dismiss self-scratching as 
unimportant: to do so would require a redefinition of tool-use in non-human
animals. Moreover, the fact that we found strong phylogenetic clustering of 
self-scratching tool use from the video survey with other examples of tool use
from the literature supports a common underlying cognitive mechanism. In line 
with this, some of the species of parrots that use tools for self-scratching in
captivity also use tools for other purposes in the wild [@Auersperg2012;
@Bastos2021;@Goodman2018;@OHara2021].

There was little evidence to suggest that the observations of self-scratching
tool use in the video survey were merely unintentional accidents or explicitly 
trained behaviours. Individual parrots often used tools slowly and repetitively 
over long periods of time, even across multiple different videos, suggesting 
that their behaviour was not random or accidental [@Sandor2020]. Parrots 
preferentially employed self-scratching tools on areas of their body that were 
otherwise inaccessible, with 96% of all instances involving scratching of the 
head or neck, suggesting intentional tool use. For 60% of the species in the 
video survey, we found two or more videos of repetitive and sustained scratching 
by different individuals of the same species in separate households, suggesting 
that the manipulations were intentional and recurring events that did not 
represent unusual stereotypies of any single individual. Regarding the 
possibility of training or cueing from owners, over half of the videos contained 
no evidence of human interaction aside from filming the behaviour. Humans only 
handed parrots their tools in two of the videos, and none of the videos featured 
owners directly rewarding tool use behaviours with food. It is unlikely that 
owners would specifically train parrots to scratch themselves with tools. 
Finally, the high levels of phylogenetic signal in our data provide strong 
evidence that the observations from our video survey reflect
biologically-endowed capacities for tool use rather than accidental or trained
behaviours, which would likely appear uniformly across the phylogeny.

While missing data imputation is becoming more common in phylogenetic 
analyses [@Debastiani2021], the important distinction between absence of 
evidence and evidence of absence has not been given as much attention. Our 
phylogenetic analysis provides one approach to this problem by distinguishing 
between true absences of tool use and absences of tool use due to a lack of 
research effort in the scientific literature or in videos. To achieve this, we 
explicitly modelled the measurement of the outcome variable along a research 
effort time series, such that species with lower research effort in the 
literature or in videos were likely to be censored. In line with our causal 
model, we also included relative brain size, feeding strategy, and phylogenetic
history as predictors of unobserved tool use. We encourage researchers to test
this model by directing future study efforts towards the parrot species with the
highest probabilities of being undetected tool-users. Future work should also
refine the causal model in Figure \@ref(fig:plotDAG) to provide more certain
estimates of tool use probabilities, either by including additional predictor
variables or modelling further causes of measurement error in the taxonomic
record, such as species abundance and geographic accessibility [@Ducatez2014].

In conclusion, we have shown that the scientific literature has insufficiently 
captured the full distribution of tool use in the parrot order. Our 
digital video survey more than doubled the number of known tool using parrot 
species from 11 to 28, and our phylogenetic model suggested that the true 
proportion of parrot tool users could be as high as 17% of all species in this 
order. These discoveries have implications for theories of the evolutionary 
drivers and origins of tool use in parrots. Beyond parrot tool use, the 
methods used in this study have the potential to be applied to other rarely 
observed behaviours, including tool use in other taxa [@Hunt2013], rhythmic 
entrainment in birds [@Benichov2016;@Patel2009;@tenCate2016;@Wilson2016],
teasing behaviours in primates [@Eckert2020], and tactical deception across all
animals [@BroJorgensen2010;@Byrne1988;@Byrne1991]. We hope that these methods
will continue to uncover a diverse array of ephemeral behaviours that have as
yet gone undetected in the scientific literature.

# Methods

## Video searches and coding

Our video search was conducted on YouTube in July 2020. Searches were conducted
manually by the first author over a month-long period, using the same IP address
and not logged in with a YouTube user account. Search terms included 
"parrot using tool" and variants (e.g., "macaw using tool", "lorikeet using 
tool", "parakeet using tool"), "tool use in parrot", "parrot tool use", "parrot 
scratching itself" (included after we found several videos demonstrating 
self-care tool use in previous searches) and equivalent terms (e.g., "parrot 
preening itself", "parrot grooming itself", "parrot scratching"). For all 
species that did not display results including object manipulation or scratching 
behaviours, we also searched the species’ common name(s) + "tool use", as well 
as the species’ common name(s) + "scratching". We also searched for translations 
of the terms "parrot tool use" and "parrot scratching" in languages for all 
countries where bird ownership was reported as >5% [@GlobalGfkSurvey], namely, 
Turkish, Czech, Polish, French, Italian, Dutch, German, Russian, Spanish, 
Portuguese, and Mandarin. Browser search histories were not cleared between
searches.

When we found a relevant video, we also searched for similar content uploaded by 
the same person/channel. For each YouTube search conducted, we watched all 
relevant videos until we reached five consecutive videos that did not feature 
any parrots. At this point, we ended that search and initiated the next search. 
In line with previous recommendations [@Nelson2013], we planned to exclude any 
videos that consisted of four or more shots edited together so as to ensure the 
behaviours being observed were not edited or manipulated, but none of the videos 
obtained qualified for exclusion.

All videos featuring parrots manipulating objects were investigated for 
potential tool use or borderline tool use. We defined tool use as the 
manipulation of an unattached object as an extension of the beak or foot to 
achieve a goal towards another object, individual, or oneself [@Shumaker2011]. 
Borderline tool use was similarly defined, except that it involved the use of an 
object that was still attached to a substrate [@Seed2010]. For example, if 
individuals used a fallen feather or stick for self-scratching this was defined 
as tool use, but using one’s currently attached tail feathers or cage 
furnishings for the same purpose was defined as borderline tool use. 
Self-scratching had to involve slow and repeated movements of touching an object 
to one’s body (or, in the case of borderline tool use, rubbing repetitively 
against an attached object [@Sandor2020]).

All relevant videos were coded for video length, species, tool use presence 
(yes/borderline), tool use type (e.g., scratching, feeding), the object being 
used (e.g., feather, stick), tool use target, human interaction 
(talking or handing object to parrot, holding parrot), and the number of shots 
within each video. Our complete dataset also includes the name for each video, 
link, subject name, sex (as declared by owner, as most parrot species are not 
sexually dimorphic), publishing date, and dates found and coded.

Since YouTube is constantly growing and evolving in both its uploaded content 
and its proprietary recommendation algorithms, the same search strategy today 
would not return exactly the same videos as our search in July 2020. It is 
possible that some videos have since been deleted or more videos of parrot tool
use have since been uploaded. However, this does not affect our conclusions
about the number of parrot species found to use tools in our initial search.

## Data for parrot species

We collected data for 194 parrot species (Supplementary Figures
\@ref(fig:plotPhylo2) - \@ref(fig:plotPhylo4)). We gathered feeding strategy 
data as a dichotomous variable ("generalist" or "specialist") from the 
EltonTraits ecological database [@Wilman2016]. As per the database, specialists 
were defined as species whose diet comprised at least 70% of a single food 
source. To calculate relative brain size, we collated data from the literature 
for all known body mass (g) and brain mass (g) values across 
parrots [@Flammer2001;@Iwaniuk2005;@Mazengenya2018;@Olkowicz2016;@Schuck2008;
@Silva2017]. For all species for which we obtained body and brain mass data, we 
calculated the encephalisation quotient (EQ) using the following 
formula [@Jerison1973]: $BrainWeight / (0.12 * BodyWeight(\frac{2}{3}))$. We 
found body mass and brain mass data for a total of 194 parrot species. This 
included all tool-using species in our video dataset, with the exception of 
three species: _Diopsittaca nobilis_, _Psittacara erythrogenys_, and _Coracopsis 
vasa_. For the latter, we used values for the closely related _Coracopsis 
nigra_. The other two species were excluded from the final dataset.

For modelling purposes, we coded research effort in both the scientific 
literature and the YouTube videos. For the scientific literature, we 
operationalised research effort as the number of papers published for each 
species' Latin name up to and including the first paper containing tool use for 
that species. If no tool use had been identified in the scientific literature 
for a species, then we coded the total number of papers published to date. We 
used the scientific database Scopus for coding the number of published papers. 
For the videos, we coded research effort as the number of search hits for each
species on YouTube. If tool use had been identified on YouTube, we estimated the
number of search hits when the first video of tool use was published on YouTube,
assuming linear growth of search hits since the inception of YouTube. If tool
use had not been identified, we used the current number of search hits.

For phylogenetic data, we used the phylogenetic tool at www.birdtree.org
[@Jetz2012] to compile 1000 posterior draws of phylogenetic trees for 174 of the 
194 parrot species for which both EQ and genomic data exist. A single maximum 
clade credibility tree was generated from these posterior draws for 
visualisation purposes. In our analyses, we iterated over posterior draws of the 
phylogeny to account for phylogenetic uncertainty.

## Phylogenetic signal

We used the _fitDiscrete_ function in the _ape_ R package [@Paradis2019] to 
calculate phylogenetic signal, for both the pre-survey and post-survey tool use 
data. We iterated the model over 100 posterior parrot phylogenies to incorporate 
phylogenetic uncertainty.

## Causal model of tool use

To infer unobserved probabilities of tool use across parrots, we proposed a 
causal model of observed tool use (Figure \@ref(fig:plotDAG)). We assumed that 
observed tool use in the scientific literature and in the videos is caused by 
both the unobserved presence or absence of tool use and research effort, proxied
by the number of papers published on a species and the number of videos
published on a species. Tool users are more likely to be observed if they are
well studied, but understudied tool users may go undetected. In addition, based
on theory, we also assumed that unobserved tool use is caused by feeding 
strategy and relative brain size [@Lefebvre2004;@Lefebvre1997;@Lefebvre2002;
@Ducatez2015;@Sol2005;@Overington2011;@HenkeVonDerMalsburg2020]. Finally, we 
assumed that shared phylogenetic history causes unobserved confounding and 
non-independence in unobserved tool use, feeding strategy, and relative brain 
size across the parrot phylogeny.

## Bayesian phylogenetic survival cure model

Given our proposed causal model, we constructed a statistical model to impute 
unobserved probabilities of tool use and test existing theories of the evolution 
of tool use in parrots. To understand the model, suppose that we have the 
following observed variables for parrot species $i$. For the scientific 
literature, we declare $\text{N}_{\text{Lit},i}$ as the number of papers 
published before and up to tool use identification for species $i$ (or, if tool 
use has not been identified, the total number of papers published for species 
$i$) and $\text{T}_{\text{Lit},i}$ as a binary variable stating whether (1) or 
not (0) tool use has yet been observed in the scientific literature for species 
$i$. For the videos, we declare $\text{N}_{\text{Vid},i}$ as the 
number of videos published before and up to tool use identification for species 
$i$ (or, if tool use has not been identified, the total number of videos 
published for species $i$) and $\text{T}_{\text{Vid},i}$ as a binary variable 
stating whether (1) or not (0) tool use has yet been observed in the 
videos for species $i$. Additionally, $\text{F}_i$ and 
$\text{EQ}_i$ are feeding strategy and encephalisation quotient values for 
species $i$ and we have a phylogenetic distance matrix $D$ that describes the 
patristic distances between all parrot species.

We assume that species $i$ is a non-tool-user with some 
probability $p_i$. We also assume that tool use is identified in the scientific 
literature and the videos at constant rates $\lambda_{\text{Lit}}$ 
and $\lambda_{\text{Vid}}$ following exponential survival functions. Given these 
assumptions, we can then describe the different ways in which variables 
$\text{N}_\text{Lit}$ and $\text{N}_\text{Vid}$ can be distributed. Focusing on 
the scientific literature, if tool use has been observed 
($\text{T}_{\text{Lit},i} = 1$), then the likelihood for 
$\text{N}_{\text{Lit},i}$ is:

\begin{align}
\text{Pr}(\text{N}_{\text{Lit},i}|\text{T}_{\text{Lit},i} = 1,p_i,\lambda_\text{Lit}) = (1-p_i)\cdot\text{Exponential}(\text{N}_{\text{Lit},i}|\lambda_\text{Lit})
\end{align}

On the other hand, if tool use has not yet been observed 
($\text{T}_{\text{Lit},i} = 0$), there are two ways that the outcome variable 
could have been realised. First, the species could be a non-tool-user with 
probability $p_i$. Second, the species could be a tool-user with probability 
($1 - p_i$) that has been censored and has not had its tool use measured yet. 
Together, then, the likelihood for $\text{N}_{\text{Lit},i}$ is:

\begin{align}
\text{Pr}(\text{N}_{\text{Lit},i}|\text{T}_{\text{Lit},i} = 0,p_i,\lambda_\text{Lit}) = p_i + ((1-p_i)\cdot\text{Exponential-CCDF}(\text{N}_{\text{Lit},i}|\lambda_\text{Lit}))
\end{align}

The Exponential-CCDF function allows for the censored nature of the data. The 
same data generating process is assumed to underlie the videos.

We define the mixture likelihood $\text{SurvivalCure}$ as the distribution 
above, with parameters $p$ (the probability of being a non-tool-user) 
and $\lambda$ (the rate of the exponential distribution). We use an 
Ornstein-Uhlenbeck Gaussian process [@McElreath2020] to model phylogenetic 
covariance. Below, we specify the full model with priors:

\begin{align}
N_{\text{Lit},i} &\sim \text{SurvivalCure}(\lambda_{\text{Lit},i}, p_{i}) \\
N_{\text{Vid},i} &\sim \text{SurvivalCure}(\lambda_{\text{Vid},i}, p_{i}) \nonumber \\
\lambda_{\text{Lit},i} &= 1 / \text{exp}(\gamma_\text{Lit}) \nonumber \\
\lambda_{\text{Vid},i} &= 1 / \text{exp}(\gamma_\text{Vid}) \nonumber \\
\text{logit}(p_{i}) &= \alpha_{\text{FEEDING}[i]} + \beta\text{EQ}_{i} +  k_{\text{SPECIES}[i]} \nonumber \\
\begin{pmatrix}k_{1}\\k_{2}\\...\\k_{n}\\\end{pmatrix}
&\sim \text{MVNormal}
\begin{pmatrix}\begin{pmatrix}0\\0\\...\\0\\\end{pmatrix},\bf{K}\end{pmatrix} \nonumber \\
\bf{K}_{ij} &= \eta^2\text{exp}(-\rho^2D_{ij}) \nonumber \\
\gamma_\text{Lit}, \gamma_\text{Vid}, \alpha_{1,2}, \beta &\sim \text{Normal(0, 1)} \nonumber \\
\eta^2, \rho^2 &\sim \text{Exponential(0.5)} \nonumber
\end{align}

The priors for this model produce reasonable prior predictions of the 
probabilities of tool use for each parrot species (Supplementary Figure
\@ref(fig:plotSurvCure7)), but a sensitivity analysis revealed that the ranking 
and posterior probabilities reported in the main text were robust to modifying 
these priors (Supplementary Figure \@ref(fig:plotSurvCure8)). We estimated the 
posterior distribution of this model using Hamiltonian Monte Carlo as 
implemented in Stan version 2.26.1 [@Stan2020]. We ran the model for 4000 
samples, with 2000 warmup samples, and iterated the model over 100 posterior 
parrot phylogenies to incorporate phylogenetic uncertainty. R-hat values and 
effective sample sizes suggested that the model converged normally. Trace plots 
are reported in Supplementary Figure \@ref(fig:plotTrace). We report 
equal-tailed credible intervals to describe the posterior distribution of this 
model in the main text.

To validate our method, we fitted the model to 100 simulated datasets with known 
parameters. The model was able to successfully recover those parameters 
(Supplementary Figure \@ref(fig:plotSurvCureSim)). We also ran a 
leave-one-species-out exercise to ensure that we could accurately predict known 
tool users. We repeated this approach for each known tool user by setting 
observed tool use to zero. Cross-validation results are reported in the main 
text.

## Ancestral state reconstruction

To determine whether the identification of novel tool-using species has 
implications for our understanding of the evolutionary origins of tool use in 
parrots, we fitted three exploratory ancestral state reconstruction models. We 
used the _ancThresh_ function from the _phytools_ R package [@phytools], 
iterating the function over 100 posterior parrot phylogenies. This function 
estimates discrete ancestral states by assuming the evolution of a latent 
continuous variable following an Ornstein-Uhlenbeck process. We fitted this 
model to three different outcome variables: (_i_) presence vs. absence of tool 
use in scientific literature only, (_ii_) presence vs. absence of tool use in 
literature and/or videos, and (_iii_) the median predicted probabilities of tool 
use from the phylogenetic survival cure model.

## Reproducibility

All analyses were conducted in R v4.2.1. [@RCoreTeam]. Visualisations were 
produced using the _ggtree_[@ggtree], _ggplot2_ [@Wickham2016], and _cowplot_ 
[@Wilke2020] packages. The manuscript was reproducibly generated using the 
_targets_ [@Landau2021] and _papaja_ [@Aust2022] packages. Code to reproduce all 
analyses and figures can be found here: 
https://github.com/ScottClaessens/phyloParrot

\newpage
\nolinenumbers

# Acknowledgements

This project was made possible through the support of a grant from the Templeton 
World Charity Foundation (A.H.T., X.J.N.). The authors would like to thank 
Daniel Sol for providing feedback on a previous version of the manuscript.

# Author Contributions

All authors contributed to the conceptualisation of the paper. A.P.M.B., X.J.N., 
and A.H.T. developed the video search methodology. S.C., D.W., and Q.A.D. 
developed the statistical models and analysed the data. All the authors wrote 
the manuscript and approved the final version for submission.

# Competing Interests

The authors declare no competing interests.

# Data Availability

All data used in this study are publicly available on GitHub: 
https://github.com/ScottClaessens/phyloParrot

# Code Availability

All code to reproduce the analyses in this study are publicly available on 
GitHub: https://github.com/ScottClaessens/phyloParrot

\newpage

# References

\begingroup

<div id="refs" custom-style="Bibliography"></div>
\endgroup

\newpage
\vspace*{60mm}

```{r child = "appendix.Rmd"}
```