### 2024-07-12 manuscript.Rmd
revise and resubmit

---
title             : "Digital video platforms and phylogenetic modelling reveal parrot tool use is not rare"
shorttitle        : "Parrot tool use"

author: 
  - name          : "Amalia P. M. Bastos"
    affiliation   : "†,1,2"
  - name          : "Scott Claessens"
    affiliation   : "*†,2"
  - name          : "Ximena J. Nelson"
    affiliation   : "3"
  - name          : "David Welch"
    affiliation   : "4"
  - name          : "Quentin D. Atkinson"
    affiliation   : "‡,2"
  - name          : "Alex H. Taylor"
    affiliation   : "‡,2,3,5,6"

affiliation:
  - id            : "1"
    institution   : "\\footnotesize Department of Psychological & Brain Sciences, Johns Hopkins University, Baltimore, MD, United States"
  - id            : "2"
    institution   : "\\footnotesize School of Psychology, University of Auckland, Auckland, New Zealand"
  - id            : "3"
    institution   : "\\footnotesize School of Biological Sciences, University of Canterbury, Christchurch, New Zealand"
  - id            : "4"
    institution   : "\\footnotesize School of Computer Science, University of Auckland, Auckland, New Zealand"
  - id            : "5"
    institution   : "\\footnotesize ICREA, Pg. Lluís Companys 23, Barcelona, Spain"
  - id            : "6"
    institution   : "\\footnotesize Institute of Neuroscience, Universitat Autònoma de Barcelona, Barcelona, Spain"

abstract: |
 Putatively rare behaviours, such as tool use, are challenging to study because absence of evidence can arise either from a species’ inability to produce the behaviour or from insufficient research effort. Here, we tackle this challenge by combining data from digital platforms and phylogenetic modelling to estimate actual rates of tool use in parrots. Videos on a social media platform revealed novel instances of tool use in 17 parrot species, more than doubling the confirmed number of tool-using parrot species from 11 (3%) to 28 (7%). Phylogenetic modelling ranked additional species that are most likely to be unobserved tool users, suggesting that between 11% and 17% of extant parrot species may be tool users. These discoveries have implications for inferences about the evolutionary drivers and origins of tool use in parrots, revealing associations with relative brain size and feeding generalism and indicating several genera where tool use was likely an ancestral trait. Overall, our findings challenge the assumption that current sampling effort captures the full distribution of putatively rare animal behaviours. Combining our sampling and analysis methods offers a fruitful approach for investigating the distribution, drivers, and origins of other rare behaviours.

keywords          : "tool use; technical intelligence; parrots; digital video platforms; phylogenetic modelling"
bibliography      : ["r-references.bib"]
floatsintext      : yes
figurelist        : no
figsintext        : yes
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
numbersections    : false
documentclass     : "apa6"
classoption       : "man, donotrepeattitle"
output            : papaja::apa6_pdf
header-includes: 
  - \usepackage{array}
  - \usepackage{caption}
  - \usepackage{graphicx}
  - \usepackage{siunitx}
  - \usepackage[normalem]{ulem}
  - \usepackage{colortbl}
  - \usepackage{multirow}
  - \usepackage{hhline}
  - \usepackage{calc}
  - \usepackage{tabularx}
  - \usepackage{threeparttable}
  - \usepackage{wrapfig}
  - \usepackage{adjustbox}
  - \usepackage{hyperref}
  - \usepackage{setspace}
  - \raggedbottom
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
  - \AtBeginEnvironment{tablenotes}{\doublespacing}
  - \captionsetup[table]{font={stretch=1,small}}
  - \captionsetup[figure]{font={stretch=1,small}}
  - \nolinenumbers
  - \note{\raggedright † These authors contributed equally to this work \par ‡ These authors contributed equally to this work \par * Correspondence concerning this article should be addressed to Scott Claessens, Level 2, Building 302, 23 Symonds Street, Auckland, New Zealand. \text{E-mail:} \href{scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}} \par This working paper has not yet been peer-reviewed.}
appendix          : "appendix.Rmd"
csl               : current-biology.csl
---

\newpage
\linenumbers

```{r include=FALSE, message=FALSE, warning=FALSE}
library(papaja)
library(targets)
library(tidyverse)
```

# Introduction

Our understanding of the evolution of animal behaviour is built on the 
assumption that we have access to sufficient data [@MacLean2012; @Lefebvre2011; 
@Logan2018]. However, this is not always the case. Data on behaviours that are 
rare, fleeting, or otherwise difficult to observe are likely to be patchy and 
incomplete [@Healy2007; @Powell2017]. Among species for which such behaviours 
have not been observed, it can be difficult to differentiate between cases in 
which the species is truly incapable of producing the behaviour and cases in 
which the species is capable of producing the behaviour but the behaviour has 
not yet been observed. Such a distinction can be critical for drawing 
conclusions about the rarity and evolution of the behaviour in question.

Comparative work on the evolution of tool use is a paradigmatic example of this 
issue. The initial discoveries of tool use in chimpanzees [@Goodall1964], 
birds [@Hunt1996], dolphins [@Smolker1997], and octopuses [@Finn2009] occurred 
decades after significant advances on other more easily measurable aspects of 
their biology. Since then, scholars have proposed a clear operational 
definition of tool use applicable to all species --- the manipulation of an 
unattached object as an extension of the animal's body to achieve a 
goal [@Shumaker2011] --- and have used the distribution of species meeting this 
definition to make various claims about the evolutionary drivers of tool use 
behaviours. For example, based on the observation that bird species with 
reported tool use tend to have larger brains, researchers have identified 
higher relative brain size as a likely precondition for tool using 
capabilities [@Audet2024;@Lefebvre2002;@Lefebvre1997;@Lefebvre2004]. These 
researchers argue that larger brains are better able to integrate visual and 
somatosensory information when innovating novel behaviours, such as tool use, 
in changing environments [@Sol2005;@Sol2009] (but see ref [@Kaplan2020]). 
Similarly, researchers have used existing reports of tool use in birds to debate
the roles of generalist versus specialist feeding strategies in driving the 
evolution of tool use, with some arguing that feeding generalists require 
technical innovations to expand their dietary niche 
[@Ducatez2015;@Overington2011;@Sol2005] and others arguing that feeding 
specialists require technical innovations for extractive foraging of specific 
foods [@HenkeVonDerMalsburg2020;@MettkeHofmann2014].

However, before we can make claims of this kind, we need to know whether current 
research effort in the literature is sufficient for robust conclusions to be 
drawn about the evolution of tool use. In fact, evidence suggests that research 
effort is often systematically biased towards particular taxonomic groups, parts 
of the world that are easy to access, and species with life history traits that 
make them easier to study, such as larger distribution ranges and population 
sizes [@Ducatez2014]. This is a crucial limitation because insufficient 
observation may lead researchers to miss true instances of tool behaviours and 
thus draw premature conclusions about the evolutionary drivers and origins of 
tool use. Researchers are keenly aware of this problem and have attempted to 
deal with it in different ways, such as controlling for the number of scientific 
papers published on different species. But previous work has not yet attempted 
to quantify and explicitly model the relationship between actual tool-using 
behaviour and what is reported in the scientific literature. If more tool-using 
species exist than previously thought, this could have important implications 
for theories of the evolutionary drivers and origins of tool use and for our 
understanding of how rare this behaviour actually is.

One potentially powerful method for quantifying actual rates of rare animal 
behaviours is by collating evidence of the behaviours from digital video 
platforms [@Nelson2013]. With millions of videos posted every day, digital 
platforms like YouTube offer a valuable source of data on animal behaviour. 
Digital video platforms have already been successfully used to uncover a variety
of rare animal behaviours, including interspecies play in dogs [@Nelson2013], 
novel problem-solving behaviours in horses [@Krueger2019], and death-related 
behaviours in Asian elephants [@Pokharel2022]. By casting the net wider than the 
scientific literature, digital video platforms can provide an indication of the 
tool-using species that the literature might be missing.

Even after collating evidence from digital video platforms, some tool-users 
could _still_ remain unobserved. One principled framework for identifying these
unobserved species is to specify a causal model of the process that generates
the observed data. We propose one such causal model in Figure 
\@ref(fig:plotDAG). In this model, we assume that the presence of tool use in 
the scientific literature (or on digital video platforms) is caused by both 
unobserved tool use capabilities and the number of published studies (or the 
number of videos) for any given species. Tool users are more likely to be 
observed if they are well studied, but understudied tool users may go 
undetected. Furthermore, based on existing theories of the evolution of tool use
[@Ducatez2015;@HenkeVonDerMalsburg2020;@Lefebvre1997;@Lefebvre2002;
@Lefebvre2004;@MettkeHofmann2014;@Overington2011;@Sol2005;@Sol2009], we propose
that the unobserved tool use capabilities are additionally caused by relative 
brain size, feeding strategy, and shared phylogenetic ancestry. Expressing this
causal model as a statistical model can suggest further species which are likely
to be unobserved tool-users and, simultaneously, test existing theories of the 
evolutionary drivers of tool use without incorrectly assuming that absence of 
evidence is evidence of absence.

(ref:plotDAGCaption) _Causal model of observed tool use._ Directed acyclic graph 
of the causal relationships between observed tool use and other variables. 
Available scientific data on tool use is caused both by unobserved tool use 
presence and scientific research effort (i.e., number of publications). 
Available video data on tool use is caused both by unobserved tool use 
presence and video research effort (i.e., number of videos). According to 
theory, unobserved tool use presence should be caused by relative brain size 
(encephalisation quotient) and feeding strategy (generalist vs. specialist). 
These variables all share unobserved confounds generated by shared phylogenetic
history. Grey circles indicate unobserved variables.

```{r plotDAG, echo=F, fig.width=6, fig.height=3.6, fig.cap="(ref:plotDAGCaption)"}
tar_read(plotDAG)
```

Here, we apply these approaches to tool use in the parrot order. We focus on 
tool use in parrots for a number of reasons. First, the scientific literature 
suggests that only a small proportion of extant parrot species (11 out of 398;
3%) use tools [@Auersperg2009;@Auersperg2010;@Auersperg2011a;@Auersperg2011b;
@Auersperg2012;@Bastos2021;@BentleyCondit2010;@Borsari2005;@Goodman2018;
@Heinsohn2017;@Janzen1976;@Lambert2015;@Lefebvre2002;@OHara2021]. Parrot tool 
use thus provides an ideal test case for examining how robust sampling is in the 
scientific literature. Second, parrots are highly popular as pets. Over 70% of 
all extant parrot species are bred in the aviculture industry and kept as pets 
worldwide [@Anderson2003;@Carrete2008;@Drews2001;@Kelly2014;@Li2014;@Su2015;
@Young2012], enabling us to leverage the power of digital video platforms to 
search for evidence of tool use [@Nelson2013]. Third, detailed data on relative
brain sizes [@Flammer2001;@Iwaniuk2005;@Mazengenya2018;@Olkowicz2016;
@Schuck2008;@Silva2017], feeding strategies [@Wilman2016], and shared ancestry
[@Jetz2012] exist for parrots, allowing us to fit the statistical model implied
by Figure \@ref(fig:plotDAG) to the entire parrot order.

We first present the results from our video survey, in which we collated videos
of tool use in parrots from a digital video platform. This survey reveals a 
number of previously unidentified tool-using parrot species, which we map onto 
the phylogeny of the parrot order. We then describe our statistical model in 
more detail, and use it to (_i_) rank further parrot species that are likely 
unobserved tool users and (_ii_) re-examine key hypotheses regarding the 
evolutionary drivers and origins of tool use in parrots.

# Results

## Digital video platform reveals tool use in additional parrot species

We surveyed the digital media platform YouTube for video evidence of tool use in 
parrots (see Methods for detailed search criteria). In our search, we used the 
standard criteria for identifying tool use in the literature, defining "true" 
tool use behaviour as the manipulation of an unattached object as an extension 
of the animal's body to achieve a goal [@Shumaker2011], while "borderline" tool 
use involved the use of an object that was still attached to a substrate
[@Seed2010].

```{r echo=F}
tar_load(crowdData)
trueCrowdData <- dplyr::filter(crowdData, ToolUse == "Y")
bordCrowdData <- dplyr::filter(crowdData, ToolUse == "B")
```

In total, we found `r nrow(crowdData)` videos of 
`r crowdData %>% pull(IndividualID) %>% unique() %>% length()` individuals from 
`r crowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
parrot species performing behaviours that met the definition of either true tool 
use (`r nrow(trueCrowdData)` videos of 
`r trueCrowdData %>% pull(IndividualID) %>% unique() %>% length()` 
individuals from 
`r trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
species) or borderline tool use (`r nrow(bordCrowdData)` videos of 
`r bordCrowdData %>% pull(IndividualID) %>% unique() %>% length()` 
individuals from 
`r bordCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
species). All videos featured pet parrots in captive settings. In 
`r crowdData %>% pull(HumanInfluence) %>% str_starts("No") %>% sum()` of these 
videos, owners did not appear to interact with the subjects. In 
`r crowdData %>% pull(HumanInfluence) %>% str_starts("Yes") %>% sum()` videos, 
there was potential human interaction, either from the owners being in close physical 
contact with the bird (e.g., bird perching on hand), talking to the bird, or 
handing it the tool (which occurred in only two videos). We could not establish 
the degree of human interaction in the remaining 
`r crowdData %>% pull(HumanInfluence) %>% str_starts("Unknown") %>% sum()` 
videos, as sound had been removed or was substituted by music. None of the 
videos featured owners directly rewarding tool use behaviours with food. All 
borderline tool use cases were excluded from further analyses.

Of the `r trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
parrot species performing true tool use, 
`r trueCrowdData %>% filter(Species != "Unknown") %>% group_by(Species, IndividualID) %>% summarise(n = n(), .groups = "drop") %>% group_by(Species) %>% summarise(check = n() > 1, .groups = "drop") %>% pull(check) %>% sum()` 
were represented in our video survey by two or more individuals over multiple 
independent observations. True tool use always involved the subject using an 
object for self-scratching (`r trueCrowdData %>% mutate(BodyPartScratched = str_detect(BodyPartScratched, "Head/Neck")) %>% pull(BodyPartScratched) %>% sum()` 
videos involved scratching the head and/or neck). The most common tool (`r sum(pull(trueCrowdData, EvaluatedTool) == "Feather")`
videos) was a moulted feather. Human-made objects (e.g., pens, spoons, pieces of 
wood, cardboard) were also common. In `r sum(readxl::read_xlsx(tar_read(crowdsourcing_file))$Directionality == "Tool to body", na.rm = TRUE)` 
videos, parrots manipulated the tool while keeping their body still, rather than 
holding the tool still and moving towards it.

According to YouTube video descriptions and owner comments, 
`r trueCrowdData %>% group_by(IndividualID, Sex) %>% summarise(.groups = "drop") %>% pull(Sex) %>% (function(x) sum(x == "Male"))` 
of the individuals performing true tool use were males and 
`r trueCrowdData %>% group_by(IndividualID, Sex) %>% summarise(.groups = "drop") %>% pull(Sex) %>% (function(x) sum(x == "Female"))` 
were females. No sex information was provided for the remaining 
`r trueCrowdData %>% group_by(IndividualID, Sex) %>% summarise(.groups = "drop") %>% pull(Sex) %>% (function(x) sum(x == "Unknown"))` 
individuals. As owners provided no information on whether sex had been 
established through genetic testing, and sexual dimorphism in parrots is 
rare [@Bercovitz1987;@Hoyo2011], we could not typically ascertain if descriptions 
were accurate. It is unclear if the disproportionately large number of males in 
the sample is a consequence of owners more readily assuming their parrots are 
male when they have not been genetically tested, owners being more likely to own 
or film male parrots, or male parrots exhibiting more true tool use behaviours 
than female parrots.

Figure \@ref(fig:plotPhylo1) maps the findings from the video survey onto a 
maximum clade credibility phylogeny for the parrot order, plotted alongside 
species previously identified in the scientific literature. Before the video 
survey, 11 parrot species (3%) had been identified as tool users in the 
scientific literature. Across our video survey, we observed true tool use in 
`r trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
species, 5 of which overlapped with the scientific literature and 
`r (trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()) - 5` 
of which were novel species. All of the species identified in the video survey 
were cockatoos (_Cacatuidae_), Old World parrots (_Psittacinae_), or neotropical 
parrots (_Arinae_). The most common species in our survey, accounting for 
`r crowdData %>% filter(Species == "Pyrrhura molinae") %>% nrow()` videos from 
`r crowdData %>% filter(Species == "Pyrrhura molinae") %>% pull(IndividualID) %>% unique() %>% length()` 
individuals, was the green-cheeked conure (_Pyrrhura molinae_). In accordance 
with the scientific literature, the video survey found no evidence of tool use 
in any species of Psittaculidae, despite this family containing some of the most 
commonly kept pet species, including lovebirds, lorikeets, and Asian parakeet 
species. Combining both the video survey and the scientific literature, we can 
thus identify 28 tool-using parrot species overall (7%), compared to the 11 
previously reported.

(ref:plotPhylo1Caption) _Results of video survey and phylogenetic survival cure
modelling mapped onto a maximum clade credibility phylogeny of the parrot order._
Orange points in the inner ring indicate species observed in the video survey,
with point size scaled by the number of videos for each species (note that three
species observed only in the video survey are not present in the phylogeny due
to a lack of genomic data: _Psittacara erythrogenys_, _Psittacus timneh_, and 
_Aratinga nenday_). Blue points in the outer ring indicate species observed in
the scientific literature, with point size scaled by the number of papers 
published on each species. Yellow species names indicate the top ten most likely
tool-using species from our phylogenetic survival cure model which were not 
observed in the scientific literature or the video survey. Total _n_ = 
`r nrow(tar_read(d))` species.

```{r plotPhylo1, echo=F, fig.width=17, fig.height=15, fig.cap="(ref:plotPhylo1Caption)"}
tar_read(plotPhylo1)
```

The identification of new tool-using species in our video survey increases the 
extent to which phylogeny can explain the distribution of tool use in the parrot 
order. We estimated phylogenetic signal (Pagel's $\lambda$) of tool use using 
both the pre-video-survey and post-video-survey data. Pagel's $\lambda$ varies 
between 0 and 1, where 0 implies that the distribution of a trait across species 
is unexplained by phylogenetic relatedness and 1 implies that the distribution 
of a trait across species is fully explained by phylogeny. Using the evidence of 
tool use from the scientific literature alone (pre-video-survey data; 11 
tool-using species), we estimated an average posterior Pagel's $\lambda$ of 
`r printnum(mean(tar_read(phySignal1)))` (95% credible interval [`r printnum(quantile(tar_read(phySignal1), 0.025))` 
`r printnum(quantile(tar_read(phySignal1), 0.975))`]; total _n_ = 
`r nrow(tar_read(d))` species). This estimate was moderate-to-strong, but highly 
uncertain. In comparison, combining the evidence from both the literature and 
the video survey (post-video-survey data; 28 tool-using species) resulted in a 
stronger and more certain estimate of phylogenetic signal. With these data, we 
estimated Pagel's $\lambda$ = `r printnum(mean(tar_read(phySignal2)))` (95% CI 
[`r printnum(quantile(tar_read(phySignal2), 0.025))` 
`r printnum(quantile(tar_read(phySignal2), 0.975))`]; total _n_ = 
`r nrow(tar_read(d))` species). Thus, the results of our video survey increase 
the extent to which the distribution of tool use across parrot species can be 
explained by shared phylogenetic ancestry. This suggests that we can potentially 
use phylogenetic information, along with other variables, to identify further 
tool-using parrot species that may remain undetected.

## Phylogenetic survival cure modelling identifies further candidate tool users

In addition to the 28 tool-using species identified in the literature and our 
video survey, we fitted a Bayesian phylogenetic survival cure model to rank 
further species that are likely to be undetected tool-users (i.e., tool-using 
species with no tool use reported in the literature or in videos).

Survival cure models [@Amico2018], also known as split population 
models [@Schmidt1989], are used to analyse the time to some event of interest 
with the added assumption that a certain proportion of the population will never 
experience the event, no matter how long they are measured for. These models 
have been used to analyse a variety of right-censored outcomes, from cancer 
relapse [@Amico2018] to criminal recidivism [@Schmidt1989]. The data are 
right-censored because some individuals will have experienced the event when 
they are measured (e.g., disease onset, return to prison) while others will have 
not experienced the event. For those who have not experienced the event, this 
may be because (_i_) the event has not happened to them yet or (_ii_) the event 
will _never_ happen to them. Survival cure models treat these two processes 
separately.

Our tool use problem has the same features. We are modelling a time-to-event; 
specifically, the amount of "time" (i.e., observation opportunities measured as 
the number of published papers or videos) until tool use is identified. This is
right-censored data because many species will not have had tool use identified
when we measure them. Moreover, we can assume that a certain proportion of the
population will never experience the event -- that is, they are a non-tool-user
and so we will never identify tool use no matter how long we measure them for.

In our model, we infer the tool-using status of each species by allowing each 
species to have their own probability of being a non-tool-user. Following our 
causal model (Figure \@ref(fig:plotDAG)), we predict these probabilities based 
on feeding strategy, encephalisation quotient, and phylogenetic history (see 
Methods for full model). The model additionally takes research effort into 
account by allowing that, among species for which tool use is unobserved, all 
else being equal those with fewer published papers and fewer video search hits 
have a higher probability of being undetected tool users (Figure 
\@ref(fig:plotSurvCure3)).

```{r echo=FALSE}
tar_load(plotSurvCure1)
```

We found that this phylogenetic survival cure model was able to adequately 
distinguish between species with and without evidence for tool use, with an 
area-under-the-curve classification statistic of 
`r printnum(tar_read(auc), digits = 2)` (Figure 
\@ref(fig:plotSurvCure9)). To further estimate the accuracy of the model's 
predictions, we also used a leave-one-species-out approach with known tool 
users. For each of the 25 tool-using species that were represented on the 
phylogeny and for which we had brain size and genomic data (we lacked data for 
three tool-using species), we fitted the model to a modified dataset which set 
tool use to be absent for the target species in both the scientific literature 
and the video survey. Across `r length(tar_read(cv))` cross-validation models, 
`r sum(tar_read(cv))` models (`r printnum(mean(tar_read(cv))*100, digits = 0)`%) 
continued to predict the target species as having a median posterior probability 
of tool use that was within the range of all other tool users. This 
classification rate was greater than the baseline classification rate of 
`r printnum(mean(plotSurvCure1$data %>% filter(ToolUse == "Absent") %>% pull(med) > 0.16478274) * 100, digits = 0)`% 
for species without evidence of tool use in the full model
(`r sum(plotSurvCure1$data %>% filter(ToolUse == "Absent") %>% pull(med) > 0.16478274)` 
of `r nrow(plotSurvCure1$data %>% filter(ToolUse == "Absent"))` species without 
evidence of tool use had a median posterior probability of tool use that was 
within the range of the tool-using species). Together, the area-under-the-curve 
statistic and the leave-one-species-out approach suggest that the model is able 
to adequately classify known tool users, with some error.

Figure \@ref(fig:plotSurvCure1) visualises the ranked posterior probabilities of 
tool use from the phylogenetic survival cure model for all parrot species. As 
expected, the known tool users are ranked towards the top of this list. However, 
several "tool use absent" species also rank highly on the list, despite not 
being identified as tool users in the scientific literature or in our video 
survey. In fact, according to the model, the most likely tool user is a species 
for which tool use is unobserved in our data: the blue-eyed cockatoo (_Cacatua 
ophthalmica_). This species is endemic to Papua New Guinea and is relatively 
understudied, with only `r tar_read(d)$EX.Pre.Tool.Papers[165]` published papers 
and `r tar_read(d)$SearchHits[165]` video search hits, which is fewer than the 
model expects are necessary to discover tool use when it is present (Figure 
\@ref(fig:plotSurvCure2)). This species is also found in the _Cacatua_ genus, a 
clade containing several known tool users. This prediction makes sense given the 
high phylogenetic signal for tool use reported by the model (
Figures \@ref(fig:plotSurvCure5) and \@ref(fig:plotSurvCure6)). Beyond the 
blue-eyed cockatoo, other highly ranked species without observed evidence of 
tool use are the Meyer's parrot (_Poicephalus meyeri_), the golden parakeet 
(_Guaruba guarouba_), the long-billed corella (_Cacatua tenuirostris_), the 
Solomons cockatoo (_Cacatua ducorpsii_), the red-fronted parrot (_Poicephalus 
gulielmi_), the Cape parrot (_Poicephalus robustus_), the yellow-eared parrot 
(_Ognorhynchus icterotis_), the red-vented cockatoo (_Cacatua haematuropygia_), 
and the gang-gang cockatoo (_Callocephalon fimbriatum_). Figure 
\@ref(fig:plotPhylo1) plots these species on the parrot phylogeny, using the top 
ten highest ranked species without observed evidence of tool use as an arbitrary 
cutoff for visualisation purposes.

(ref:plotSurvCure1Caption) _Posterior predicted probabilities of tool use for 
each species from our phylogenetic survival cure model._ Points are posterior 
medians and lines are 50% and 95% credible intervals. Total _n_ = 
`r nrow(tar_read(d))` species.

```{r plotSurvCure1, echo=F, fig.width=7, fig.height=10, fig.cap="(ref:plotSurvCure1Caption)"}
tar_read(plotSurvCure1)
```

(ref:plotSurvCure2Caption) _Expected number of published papers and videos until 
tool use discovery, according to the survival component of the phylogenetic 
survival cure model._ Densities are full posterior distributions, points are 
posterior medians, and lines are 50% and 95% credible intervals.

```{r plotSurvCure2, echo=F, fig.width=7, fig.height=3, fig.cap="(ref:plotSurvCure2Caption)"}
tar_read(plotSurvCure2)
```

```{r echo=F}
tar_load(sumProbs)
```

The posterior probabilities shown in Figure \@ref(fig:plotSurvCure1) are 
estimated with uncertainty, so it is difficult to "identify" any particular 
species as an undetected tool user. Nevertheless, taking the sum of all the 
posterior probabilities for the 
`r nrow(plotSurvCure1$data %>% filter(ToolUse == "Absent"))` species without 
observed evidence of tool use, we can estimate that around 
`r printnum(median(sumProbs), digits = 0)` of those species are likely to be 
undetected tool users (median sum of probabilities = 
`r printnum(median(sumProbs), digits = 2)`, 95% CI
[`r printnum(quantile(sumProbs, 0.025), digits = 2)` 
`r printnum(quantile(sumProbs, 0.975), digits = 2)`]). When combined with the 
species known to use tools, this implies that between 11% and 17% of extant 
parrot species may be tool users.

## Implications for the evolutionary drivers and origins of tool use

```{r echo=FALSE}
tar_load(postFull)
diff <- postFull$aP[,2] - postFull$aP[,1]
```

The predicted probabilities from our phylogenetic survival cure model have 
implications for inferences about the evolutionary drivers and origins of tool 
use in the parrot order. Regarding the drivers of tool use hypothesised in 
Figure \@ref(fig:plotDAG), the phylogenetic survival cure model revealed that 
encephalisation quotient strongly positively predicted the probability of tool 
use (median posterior log odds slope = `r printnum(median(-postFull$bEQ))`, 95% 
CI [`r printnum(quantile(-postFull$bEQ, 0.025))` 
`r printnum(quantile(-postFull$bEQ, 0.975))`]; total _n_ = `r nrow(tar_read(d))` 
species; Figure \@ref(fig:plotSurvCure4)). This helps explain the ranking in 
Figure \@ref(fig:plotSurvCure1): the blue-eyed cockatoo has the largest relative 
brain size in the dataset. We also found that feeding generalist species were 
slightly more likely to be tool users, though the posterior difference between 
generalists and specialists was quite uncertain (median posterior log odds 
difference = `r printnum(median(diff))`, 95% CI [
`r printnum(quantile(diff, 0.025))` `r printnum(quantile(diff, 0.975))`]; total 
_n_ = `r nrow(tar_read(d))` species). These results from the survival cure model 
differed from the results of models fitted to pre-video-survey and 
post-video-survey data without the survival cure component, which found no 
effect of relative brain size and no difference between feeding strategies, 
respectively (Figure \@ref(fig:plotComparison)).

(ref:plotSurvCure4Caption) _Posterior predictions for the effects of feeding 
strategy and encephalisation quotient on the probability of tool use from the 
phylogenetic survival cure model._ In the left plot, points and lines represent 
posterior medians and 50% and 95% credible intervals, with densities 
representing full posterior distributions. In the right plot, the line and 
shaded areas represent the posterior median regression line with 50% and 95% 
credible intervals. In both plots, individual species are coloured according to 
the presence / absence of tool use in the video survey and the scientific 
literature. Total _n_ = `r nrow(tar_read(d))` species, generalist _n_ = 
`r sum(tar_read(d)$Feeding == "Generalist")` species, specialist _n_ = 
`r sum(tar_read(d)$Feeding == "Specialist")` species.

```{r plotSurvCure4, echo=F, fig.width=8, fig.height=4, fig.cap="(ref:plotSurvCure4Caption)"}
tar_read(plotSurvCure4)
```

Regarding the origins of tool use, we fitted exploratory ancestral state 
reconstruction models to the pre-video-survey data, the post-video-survey data, 
and the predicted probabilities from the phylogenetic survival cure model. The 
discoveries from our video survey and from our phylogenetic modelling increased 
the likelihood that the capacity for tool use was present in the most recent 
common ancestors for several parrot genera. These include the most recent common 
ancestor of amazon parrots native to the Americas (_Amazona_), the most recent 
common ancestor of the true white cockatoos and corellas found in South East 
Asia and Australasia (_Cacatua_), the most recent common ancestor of the kea and 
the kākā from New Zealand (_Nestor_), and the most recent common ancestor for
the _Poicephalus_ genus native to Africa (Table \@ref(tab:tableASR); 
Figures \@ref(fig:plotASR1) - \@ref(fig:plotASR3)).

(ref:tableASRCaption) Estimated probabilities of tool use for most recent common 
ancestors of several parrot genera. _Probabilities estimated using exploratory 
ancestral state reconstruction models fitted to the pre-video-survey data, 
post-video-survey data, and predicted probabilities from the phylogenetic 
survival cure model. Total_ n _= `r nrow(tar_read(d))` species._

```{r tableASR, echo=F, warning=F, message=F}
apa_table(
  tar_read(tableASR),
  caption = "(ref:tableASRCaption)",
  escape = FALSE
)
```

# Discussion

Since the earliest anecdotes of parrots using tools by Wallace in the 1880s
[@Wallace1869] and more systematic anecdotal reports in the 1970s [@Goodall1971], 
only 11 parrot species (3% of all extant parrots) have been documented as tool 
users in the scientific literature. Our study collated data from a digital video
platform to identify 17 additional tool-using parrot species that are new to 
science, more than doubling the overall count to 28 species (7%). These species 
consisted of cockatoos (_Cacatuidae_), Old World parrots (_Psittacinae_), and 
neotropical parrots (_Arinae_). Beyond the video survey, the strong 
phylogenetic signal in our dataset allowed us to use phylogenetic information, 
along with other variables, to infer the unobserved probabilities of tool use 
across the parrot order. Our phylogenetic survival cure model incorporated
information on phylogenetic history, research effort, relative brain size, and 
feeding specialisation to rank parrot species that were most likely to be 
undetected tool users. The sum of probabilities from this model implied that 
between 15 and 41 of the species without observed evidence of tool use are 
likely to be undetected tool users, suggesting that the true proportion of tool
users may be as high as 17%.

These findings have a number of important implications. First, our findings show 
that current research effort in the scientific literature is insufficient to 
capture the real world occurrence of parrot tool use. If the scientific 
literature had sampled the natural world sufficiently, we would expect to see 
close correspondence between those species reported as tool users in the 
literature and those species the public have filmed performing tool use. 
Instead, we discovered a large discrepancy between these two data sources, both 
in the prevalence of tool use and the species identified, likely due to the 
numerous difficulties inherent to conducting observations in the wild. This 
raises the possibility that other rare behaviours remain yet to be discovered in 
the scientific literature.

Second, in terms of the evolution of tool use in parrots, our study challenges a 
key assumption made in the literature to date: that only a minority of parrots 
are tool users [@Auersperg2012;@Bastos2021;@Goodman2018;@Lambert2015;@OHara2021]. 
The paucity of evidence for tool use across parrots in the literature initially 
implied that the capacity for tool use may have evolved independently in 
different parrot species. Our discovery of the widespread distribution of tool 
use across the parrot phylogeny, along with the strong phylogenetic signal in 
this expanded dataset, challenges this and suggests that, at least for some 
parrot clades, the capacity for tool use might be a homologous trait that has 
been evolutionarily conserved. Our exploratory ancestral state reconstruction 
analysis provides preliminary support for this hypothesis, revealing an 
increased probability of the capacity for tool use among the most recent common 
ancestors for the _Amazona_, _Cacatua_, _Nestor_, and _Poicephalus_ genera. The 
complexity of tool use behaviour that such a basic capacity might afford, and 
the conditions under which it emerges in the wild, are yet to be established. 
However, even at this preliminary stage, our analysis raises an alternative 
hypothesis for the observed tool use in _Cacatua_ [@Auersperg2012;@OHara2021] 
and _Nestor_ [@Auersperg2009;@Auersperg2010;@Auersperg2011a;@Auersperg2011b;
@Bastos2021;@Goodman2018], namely that tool behaviours have arisen due to the 
common ancestor having the capacity to use tools, rather than from independent 
evolution or behavioural innovation within species.

Third, our results support existing theories of the drivers of tool use. We 
found that encephalisation was strongly positively related to the probability of 
tool use in our phylogenetic model, supporting previous theories linking 
relative brain size to increased tool innovation in birds [@Lefebvre1997;
@Lefebvre2002;@Lefebvre2004] and primates [@Reader2002]. We acknowledge that
encephalisation quotient is not a perfect measure due to measurement error and
challenges with interpretation [@Healy2007;@Logan2018]. However, encephalisation
quotient has much greater coverage across the parrot phylogeny than more 
fine-grained measures like whole neuron count [@Sol2022] and there is no reason 
to think that measurement error would produce the consistent patterns across our
study and prior work. To understand the causal mechanisms responsible for these
relationships, we encourage further work on the specific neural correlates of 
technical intelligence in parrots, e.g., [@Alvarez2020]. In our phylogenetic 
model, we also found that tool use was somewhat more likely among feeding 
generalists compared to feeding specialists, although this difference was 
uncertain. This trend supports previous suggestions that increased cognitive 
abilities and technical innovation rates are required to expand a generalist 
species' dietary niche [@HenkeVonDerMalsburg2020;@HenkeVonDerMalsburg2021;
@Ducatez2015]. However, the trend contradicts theories linking tool use to 
dietary specialisation, whereby species eating specific foods that require 
extractive foraging have higher cognitive ability and are especially prone to 
using tools [@HenkeVonDerMalsburg2020].

All instances of tool use in our video survey met the established criteria for 
tool use in the literature: the manipulation of an unattached object as an 
extension of the animal's body to achieve a goal [@Shumaker2011]. However, a 
striking feature of the dataset is that the 17 additional species identified in
the video survey were exclusively observed using tools for self-scratching.
While parrots in the wild use tools to achieve a variety of goals, including 
extractive foraging and preening [@Goodman2018;@OHara2021], parrots in captivity
only require tools to achieve the latter goal. It could be argued that 
self-scratching is a more cognitively simplistic form of tool use than 
extractive foraging as it is egocentric, meaning that the tool is directed 
towards oneself rather than an external object. Nevertheless, self-scratching 
meets the definition for a more complex type of embodied tool use known as 
'tooling', the deliberate generation of a mechanical interface by using an 
object to manipulate another target or surface [@Fragaszy2018]. In our dataset, 
most self-scratching tooling instances involved manipulating the tool while 
keeping the body still, suggesting that these were goal-directed, deliberate 
movements by individuals. This makes it difficult to dismiss self-scratching as 
unimportant: to do so would require a redefinition of tool-use and tooling in 
non-human animals. Moreover, the fact that we found strong phylogenetic 
clustering of self-scratching tool use from the video survey with other examples 
of tool use from the literature supports a common underlying cognitive 
mechanism. In line with this, some of the species of parrots that use tools for 
self-scratching in captivity also use tools for other purposes in the wild 
[@Auersperg2012;@Bastos2021;@Goodman2018;@OHara2021].

A potential concern with data originating from digital video platforms is its 
reliability [@Auersperg2020;@DechaumeMoncharmont2020;@Farrar2020;@Nelson2013;@Sandor2020;@vonBayern2020]. 
In particular, the self-care tool use instances detected in our video survey 
might be attributable to training or unintentional cueing by the birds' owners.
However, there was little evidence to suggest that the observations of 
self-scratching tool use in the video survey were merely unintentional accidents 
or explicitly trained behaviours. Individual parrots often used tools slowly and 
repetitively over long periods of time, even across multiple different videos, 
suggesting that their behaviour was not random or accidental [@Sandor2020]. 
Parrots preferentially employed self-scratching tools on areas of their body 
that were otherwise inaccessible, with 96% of all instances involving scratching 
of the head or neck, suggesting intentional tool use. For 60% of the species in 
the video survey, we found two or more videos of repetitive and sustained 
scratching by different individuals of the same species in separate households, 
suggesting that the manipulations were intentional and recurring events that did 
not represent unusual stereotypies of any single individual. Regarding the 
possibility of training or cueing from owners, over half of the videos contained 
no evidence of human interaction aside from filming the behaviour. Humans only 
handed parrots their tools in two of the videos, and none of the videos featured 
owners directly rewarding tool use behaviours with food. It is unlikely that 
owners would specifically train parrots to scratch themselves with tools. 
Finally, the high levels of phylogenetic signal in our data provide strong 
evidence that the observations from our video survey reflect
biologically-endowed capacities for tool use rather than accidental or trained
behaviours, which would likely appear uniformly across the phylogeny. 
Nevertheless, future work using data such as this should choose target 
behaviours carefully and examine videos for evidence of training or cueing.

Our data reveals that a wide range of parrot species exhibit self-scratching 
tool use behaviours in captive environments. It is possible that these self-care 
behaviours emerge more often in captivity because pet parrots are kept with few 
or no conspecifics, and therefore must innovate tool use to preen inaccessible 
body areas, such as their heads and necks, which would have been otherwise 
preened by flock members. This may indicate that human interaction and 
allopreening in the absence of a conspecific is insufficient for pet parrots, as 
suggested by studies on welfare and chronic stress in lone captive parrots 
[@Aydinonat2014;@Engebretson2006;@Mason2010;@Meehan2003]. Beyond social 
isolation in captivity, other factors that reduce preening efficacy by an 
individual might also lead to the innovation of self-care tooling: for example,
a captive kea housed with conspecifics in a naturalistic environment was shown
to have innovated the use of a pebble tool for self-care likely as a consequence 
of a missing upper mandible [@Bastos2021]. These behaviours might therefore be 
innovated by individuals when ecologically necessary. Critically, however, a 
cognitive capacity to innovate a behaviour such as tool use must exist in a 
species in order to manifest in any environment, whether that be captive or in 
the wild. As such, despite differences in individuals' life experiences in 
captive and wild settings, such behaviour emerging in a captive setting at the 
very least indicates an underlying cognitive capacity to innovate that same 
behaviour in the wild.

While missing data imputation is becoming more common in phylogenetic 
analyses [@Debastiani2021], the important distinction between absence of 
evidence and evidence of absence has not been given as much attention. Our 
phylogenetic analysis provides one approach to this problem by distinguishing 
between true absences of tool use and absences of tool use due to a lack of 
research effort in the scientific literature or in videos. To achieve this, we 
explicitly modelled the measurement of the outcome variable along a research 
effort time series, such that species with lower research effort in the 
literature or in videos were likely to be censored. In line with our causal 
model, we also included relative brain size, feeding strategy, and phylogenetic
history as predictors of unobserved tool use. We encourage researchers to test
this model by directing future study efforts towards the parrot species with the
highest probabilities of being undetected tool-users. Future work should also
refine the causal model in Figure \@ref(fig:plotDAG) to provide more certain
estimates of tool use probabilities, either by including additional predictor
variables or modelling further causes of measurement error in the taxonomic
record, such as species abundance and geographic accessibility [@Ducatez2014].

In conclusion, we have shown that the scientific literature has insufficiently 
captured the full distribution of tool use in the parrot order. Our 
digital video survey more than doubled the number of known tool using parrot 
species from 11 to 28, and our phylogenetic model suggested that the true 
proportion of parrot tool users could be as high as 17% of all species in this 
order. These discoveries have implications for theories of the evolutionary 
drivers and origins of tool use in parrots. Beyond parrot tool use, the 
methods used in this study have the potential to be applied to other rarely 
observed behaviours, including tool use in other taxa [@Hunt2013], rhythmic 
entrainment in birds [@Benichov2016;@Patel2009;@tenCate2016;@Wilson2016],
teasing behaviours in primates [@Eckert2020], and tactical deception across all
animals [@BroJorgensen2010;@Byrne1988;@Byrne1991]. We hope that these methods
will continue to uncover a diverse array of ephemeral behaviours that have as
yet gone undetected in the scientific literature.

\newpage
\nolinenumbers

# Acknowledgements

This project was made possible through the support of a grant from the Templeton 
World Charity Foundation (A.H.T., X.J.N.). The authors would like to thank 
Daniel Sol for providing feedback on a previous version of the manuscript.

# Author Contributions

All authors contributed to the conceptualisation of the paper. A.P.M.B., X.J.N., 
and A.H.T. developed the video search methodology. S.C., D.W., and Q.A.D. 
developed the statistical models and analysed the data. All the authors wrote 
the manuscript and approved the final version for submission.

# Declaration of Interests

The authors declare no competing interests.

\newpage

# References

\begingroup

<div id="refs" custom-style="Bibliography"></div>
\endgroup

\newpage
\linenumbers

```{r child = "methods.Rmd"}
```

\newpage
\nolinenumbers
\vspace*{60mm}

```{r child = "appendix.Rmd"}
```